---
title: "Getting started with catchr"
author: "Zachary Burchill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with catchr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)


# To-do: change %\VignetteIndexEntry{Vignette Title}




```

Compared to many other programming languages, the way `R` handles 'conditions' --- errors, warnings, messages, and most things referred to as 'exceptions' in other languages --- is pretty unintuitive.

Its help file on the subject^[Accessible via `help(conditions)`.]  is basically impenetrable, and even though Hadley Wickham has stepped in and produced [a very helpful guide on conditions](http://adv-r.had.co.nz/Exceptions-Debugging.html), understanding what's going on is still pretty hard, and actually _handling_ conditions is even harder.

At its heart, `catchr` is designed to ~~make it so that you never have to read any of that stuff~~ make dealing with conditions easy from the get-go. 

## Make your own catching functions

Suppose I want to run some code, and I want it to catch any warnings it produces, "hide" them so they don't appear, and instead just store them so I can do something else with them later.  All without making my code halt or restart.

```{r, eval=FALSE}
library(catchr)
results <- catch_expr(abitrary_code, warning = c(collect, muffle))
```

Boom, done.

But let's say I want to be a good programmer and not retype the same code over and over again---let's say I want to make a _function_ that will always do that for me.

```{r, eval=FALSE}
collect_warnings_fn <- make_catch_fun(warning = c(collect, muffle))
```

One line.

But let's say ~~I'm a _naaaaasty_ boy~~^[Ed.: you can't say that in a CRAN package] I have a complex situation which necessitates printing "oops" when the code raises a message and immediately exiting it, turning any raised warnings into messages, and muffling any errors.

```{r, eval=FALSE}
weirdo_catcher <- make_catch_fun(
  message = c(function(x) print("oops"), exit),
  warning = c(tomessage, muffle),
  error = muffle
)
```

\*Shiver\*. I can do that too.



Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
